---
title: "Analysis script"
author: "Quinn Moon"
date: "03/31/2023"
output: html_document
---

This Quarto file loads the cleaned Palmer Penguins data and does some initial analysis.
The results were created by the `statistical_analysis.R` script and to the `Results` folder.

For this analysis, we want to ask and investigate three seperate questions relating to our penguin data set. This analysis will walk through an exploration of the various relationships between sex, mass, species, and flipper length. I have included some of my thinking as I worked through creating tables and figures. 

# Setup


```{r }
#| cache = FALSE
knitr::read_chunk('statistical_analysis.R')
```

Load needed packages and filepaths.  

```{r setup}
#| message: FALSE
#| warning: FALSE
```

Load custom functions.

```{r functions}
```

Load the data.

```{r loaddata}
```

# Create a summary table

Here we will use skimr which produces a summary table, but with more than we want. Letʻs save it as an object so we can grab the elements we want. 

It turns out that the `skim()` function returns a tibble. We can coerce it to a dataframe and use `head()` to see what the column names are and what the data look like. 

```{r, summarize}
```

## Select columns for a final summary table

Letʻs retain only the variable, N, mean, sd, and category counts, but letʻs rename them. Since we will use the SD to compute the stadard error letʻs just rename it as SE now and save a step later.  (Remember the standard error is just the standard deviation/ square root(sample size)). 

Note: The `results="asis"` is an option for `knitr` to layout the the results without the code formatting. Normally output from code chunks has the look of R output. `"asis"` removes that behavior, so we can lay it out as a typeset table for print. 

```{r summary.table, results="asis"}
```


# Do sexes differ in size?

First try a boxplot of mass by sex

```{r mass_sex_boxplot}
```


Now letʻs try a density plot

```{r mass_sex_density}
```

When we plot the results, Males appear to be larger than females. To test the statistical significance of this result we can use ANOVA. We see that sexes do differ in size and this is a very robust statistical result. 

```{r mass_sex_aov}
```

# Does species have an affect on size?

Based on our initial screening of the relationship between size and sex, we see that within each sex, the distribution is non-normal.
For each Sex distinction, we grouped samples of the same sex but different species together. Perhaps, the underlying assumption of species
is contributing to our non normal results. 

What about a species affect? Since we know that sexes differ, we can include a two-way ANOVA and test both factors. 


When we add species to the model _after_ we account for sex, it is highly significant. 

```{r mass_species_island_aov}
```

If we add sex _first, before_ species, it is highly signficant! We interpret this as species differing in mass as well.
```{r mass_sex_species_aov}
```
Letʻs plot both factors together to see whatʻs going on. We see that for each species, males are generally larger and there is some level of sexual dimorphism for all three species. By graphing the data by species, we can also notice that Gentoo is the largest of the three species.

```{r mass_species_sex_density}
```

After some analysis, we see that sex and species both influence mass. But, when mass changes, do other physical measurments, like flipper length also change?

# Let's adress the question, does mass have an affect on flipper length?

We include a one way ANOVA to test
```{r mass_flipper_aov}
```
From out ANOVA, we can see that the positive relationship between mass and flipper length is highly significant.
Let's graph the relationship for a better look.

```{r mass_flipper_plot}
```
We can add a linear trend line to better highlight the positive relationship between mass and flipper length that appears to increase relativly linearly. 

# Discussion

Overall, we see that both species and sex significantly impact mass while mass has a significant influence on flipper length.

# The End
